Characterizing Water Composition with an Autonomous Robotic Team Employing Comprehensive In-Situ Sensing, Hyperspectral Imaging, Machine Learning, and Conformal Prediction

### 2024-03-05 paper/paper.tex
Updates from Overleaf
final edits

\begin{figure}[t!]
\centering
\vspace{-0.3in}
\includegraphics[width=0.9\columnwidth]{paper/figures/results/fits/physical-fitres.png}
\vspace{-0.2in}
\caption{Scatter diagrams (left) and quantile-quantile plots (right) for the hyperparameter optimized RFR models for the physical variables measured by the USV.\label{fig:physical-fit}}
\end{figure}

@@

\begin{figure}[t!]
\centering
\vspace{-0.3in}
\includegraphics[width=0.9\columnwidth]{paper/figures/results/maps/physical.pdf}
\vspace{-0.2in}
\caption{Maps generated by applying each of the physical variable models to the hyperspectral data cubes collected on 11-23. Overlaid over the predictions are color-filled squares showing the associated in-situ reference data for the same collection period. The size of the squares has been exaggerated for the visualization. We note that there is good agreement between the model predictions and the reference data. \label{fig:map-physical}}
\end{figure}

@@

\begin{figure}[t!]
\centering
\vspace{-0.3in}
\includegraphics[width=\columnwidth]{paper/figures/results/fits/chemical-fitres.png}
\vspace{-0.25in}
\caption{Scatter diagrams (left) and quantile-quantile plots (right) for the hyperparameter optimized RFR models for the chemical variables measured by the USV.\label{fig:chemicals-fit}}
\end{figure}

\begin{figure}[t!]
\centering
\vspace{-0.25in}
\includegraphics[width=\columnwidth]{paper/figures/results/fits/chemical-ranking.pdf}
\vspace{-0.25in}
\caption{Ranked permutation importance for for the top 25 features of the chemical models. The permutation importance measures the deacrease in model $R^2$ value after replacing each feature in the prediction set with a random permutation of its values.\label{fig:chemicals-fi}}
\end{figure}

@@

In recent years, much effort has been spent on the curation of comprehensive datasets combining water quality records with decades of satellite imagery to enable the development of new methods for retrieving water quality parameters. For example, Aurin et al. curated over 30 years of oceanographic field campaign data with associated coincident satellite imagery \cite{aurin2018remote}. Similarly, Ross et al. combined more than 600,000 records of dissolved organic carbon, chlorophyll-a and other water quality variables with historical Landsat reflectance data for the period 1984--2019 \cite{ross2019aquasat}. The sensing paradigm we have demonstrated here was able to rapidly collect comparable volumes of data within the span of just three observation periods. Therefore, despite the fact that individual UAV tracks cover far less spatial extent than remote sensing imagery, the ability to collect coordinated in-situ measurements together with detailed hyperspectral images offers a significant improvement over these traditional approaches. With a coordinated robot team, one does not need to rely on infrequent satellite overpasses when planning data collection. Further, the time offset between reference measurements and remote sensing is significantly reduced from days to minutes.

@@

The second contribution of this study is to explore the breadth of possible water quality and composition parameters that can be accurately mapped by hyperspectral imagery collected by the UAV. The results presented here confirm the ability of the robot team to predict optically active parameters including blue-green algae, chlorophyll-a, CDOM, crude oil, optical brighteners, turbidity, and temperature. Additionally, we are also also able to infer the distributions of optically inactive variables including conductivity, pH, and ion concentrations. Other studies using multispectral and hyperspectral remote sensing imagery have also estimated optically inactive water quality parameters with the ability to do so stemming from the relationship of these variables to optically active properties of the water \cite{vakili2020determination,guo2021machine,niu2021deep}. We note that in our investigation, the models trained for many of these variables often outperformed their optically active counterparts. As the abundance of these variables is likely tied to the specific composition and content of the pond, it is unlikely that models trained for these optically inactive variables can generalized to other bodies of water.

@@

In this study, we address two key limitations of current remote sensing approaches to characterizing water quality, namely the limited spatial, spectral, and temporal resolution provided by existing satellite platforms and the lack of comprehensive in-situ measurements needed to validate remote sensing data products. By equipping an autonomous USV with a suite of reference sensors, we rapidly collect significantly more data than existing approaches which rely on the collection of individual samples for lab analysis or are constrained to continuous sensing at fixed sites. Utilizing an autonomous UAV equipped with a hyperspectral imager in tandem with the USV allows us to quickly generate aligned datasets which are used to train machine learning models mapping measured reflectance spectra to the desired water quality variables. By virtue of this increased data volume, we are able to simultaneously estimate the uncertainty of our models by using conformal prediction. Finally, the hyperspectral data cube processing pipeline employed on board the UAV makes it possible to deploy these trained models to swiftly generate maps of the target variables across bodies of water. The rapid turnaround time from data collection to model deployment is critical for real-time water quality evaluation and risk assessment.

>>>>

\begin{figure}[t!]
\centering
\vspace{-0.3in}
\includegraphics[width=0.9\columnwidth]{paper/figures/results/fits/physical-fitres.png}
\vspace{-0.1in}
\caption{Scatter diagrams (left) and quantile-quantile plots (right) for the hyperparameter optimized RFR models for the physical variables measured by the USV.\label{fig:physical-fit}}
\end{figure}

@@

\begin{figure}[t!]
\centering
\vspace{-0.3in}
\includegraphics[width=0.9\columnwidth]{paper/figures/results/maps/physical.pdf}
\vspace{-0.1in}
\caption{Maps generated by applying each of the physical variable models to the hyperspectral data cubes collected on 11-23. Overlaid over the predictions are color-filled squares showing the associated in-situ reference data for the same collection period. The size of the squares has been exaggerated for the visualization. We note that there is good agreement between the model predictions and the reference data. \label{fig:map-physical}}
\end{figure}

@@

\begin{figure}[t!]
\centering
\vspace{-0.31in}
\includegraphics[width=\columnwidth]{paper/figures/results/fits/chemical-fitres.png}
\vspace{-0.31in}
\caption{Scatter diagrams (left) and quantile-quantile plots (right) for the hyperparameter optimized RFR models for the chemical variables measured by the USV.\label{fig:chemicals-fit}}
\end{figure}

\begin{figure}[b!]
\centering
\vspace{-0.25in}
\includegraphics[width=\columnwidth]{paper/figures/results/fits/chemical-ranking.pdf}
\vspace{-0.25in}
\caption{Ranked permutation importance for for the top 25 features of the chemical models. The permutation importance measures the deacrease in model $R^2$ value after replacing each feature in the prediction set with a random permutation of its values.\label{fig:chemicals-fi}}
\end{figure}

@@

In recent years, much effort has been spent on the curation of comprehensive datasets combining water quality records with decades of satellite imagery to enable the development of new methods for retrieving water quality parameters. For example, Aurin et al. curated over 30 years of oceanographic field campaign data with associated coincident satellite imagery \cite{aurin2018remote}. Similarly, Ross et al. combined more than 600,000 records of dissolved organic carbon, chlorophyll-a and other water quality variables with historical Landsat reflectance data for the period 1984--2019 \cite{ross2019aquasat}. The sensing paradigm we have demonstrated here was able to rapidly collect comparable volumes of data within the span of just three observation periods. Therefore, despite the fact that individual UAV tracks cover far less spatial extent than remote sensing imagery, the ability to collect coordinated in-situ measurements together with detailed hyperspectral images offers a significant improvement over these traditional approaches. With a coordinated robot team, one does not need to rely on infrequent satellite overpasses when planning data collection. Furthermore, the time offset between reference measurements and remote sensing is significantly reduced from days to minutes.

@@

The second contribution of this study is to explore the breadth of possible water quality and composition parameters that can be accurately mapped by hyperspectral imagery collected by the UAV. The results presented here confirm the ability of the robot team to predict optically active parameters including blue-green algae, chlorophyll-a, CDOM, crude oil, optical brighteners, turbidity, and temperature. Additionally, we are also also able to infer the distributions of optically inactive variables including conductivity, pH, and ion concentrations. Other studies using multispectral and hyperspectral remote sensing imagery have also estimated optically inactive water quality parameters with the ability to do so stemming from the relationship of these variables to optically active properties of the water \cite{vakili2020determination,guo2021machine,niu2021deep}. We note that in our investigation, the models trained for many of these variables outperformed their optically active counterparts. As the abundance of these variables is likely tied to the specific composition and content of the pond, it is unlikely that models trained for these optically inactive variables will generalize to other bodies of water.

@@

In this study, we address two key limitations of current remote sensing approaches to characterizing water quality, namely the limited spatial, spectral, and temporal resolution provided by existing satellite platforms and the lack of comprehensive in-situ measurements needed to validate remote sensing data products. By equipping an autonomous USV with a suite of reference sensors, we rapidly collect significantly more data than existing approaches which rely on the collection of individual samples for lab analysis or are constrained to continuous sensing at fixed sites. Utilizing an autonomous UAV equipped with a hyperspectral imager in tandem with the USV allows us to quickly generate aligned datasets which are used to train machine learning models mapping measured reflectance spectra to the desired water quality variables. By virtue of this increased data volume, we are able to simultaneously estimate the uncertainty of our models by using conformal prediction. Finally, the hyperspectral data cube processing workflow employed on board the UAV makes it possible to deploy these trained models to swiftly generate maps of the target variables across bodies of water. The rapid turnaround time from data collection to model deployment is critical for real-time water quality evaluation and risk assessment.
